
---
title: "PHASE 2 A"
author: "Nourhan Berjawi"
date: "11/15/2020"
output: html_document
---

```{r setup, echo = FALSE}
setwd("~/Desktop/historypm2.5")
library(tidyverse)
library(caret) 
library(plyr)
library(dplyr) 
library(anomalize)
library(bnstruct)
library(forecast)
```

We have used the data for the PM2.5 concentrations from the past 20 years for around 900 US counties.

The first step is to read the .csv files for the years 2000 till 2020. For every year, there are two data-sets. The first one is named "pm2.5X.csv" and refers to recordings taken using the FRM/FEM method; the second one is named "pm2.5nX.csv" and refers to recordings taken using a non-FRM/FEM method.
The two datasets have values for distinct dates. This is why we decided to consider both methods, and collect as much recordings of pm2.5/county as we can.

Moreover, upon reading each data-set, we do some alterations:
  1) remove all columns except those of pm2.5 daily average, state, county, longitude, and latitude
  2) create location column
  3) order by location and date
  4) replace negative PM2.5 values with zero (according to the data-set source, the monitors record a negative value when the PM2.5 value recorded is very low and the monitor hasn't been calibrated for a while)
  5) bind the FRM/FEM data with the non-FRM/FEM data
```{r getPM2.5}

setwd("~/Desktop/historypm2.5")
pm2.52000 <- read.csv("pm2.52000.csv")
pm2.5n2000 <- read.csv("pm2.5n2000.csv")


pm2.52000 <- pm2.52000[c(6,7,12,17,25,26)]
pm2.5n2000 <- pm2.5n2000[c(6,7,12,17,25,26)]
pm2.52000$Date.Local <- as.Date(pm2.52000$Date.Local)
pm2.5n2000$Date.Local <- as.Date(pm2.5n2000$Date.Local)
pm2.52000$location <-paste(as.character(pm2.52000$State.Name),as.character(pm2.52000$County.Name))
pm2.52000$Longitude <- as.integer(pm2.52000$Longitude)
pm2.52000$Latitude <- as.integer(pm2.52000$Latitude)
pm2.5n2000$Longitude <- as.integer(pm2.5n2000$Longitude)
pm2.5n2000$Latitude <- as.integer(pm2.5n2000$Latitude)
pm2.5n2000$location <-paste(as.character(pm2.5n2000$State.Name),as.character(pm2.5n2000$County.Name))
pm2.52000 <- pm2.52000[c(1,2,3,4,7)]
pm2.5n2000 <- pm2.5n2000[c(1,2,3,4,7)]
names(pm2.5n2000)[names(pm2.5n2000) == "Arithmetic.Mean"] <- "Arithmetic.Mean2"
pm2.52000<- merge(pm2.5n2000,pm2.52000 ,by = c("location","Date.Local","Longitude","Latitude"),all = TRUE)
pm2.52000$Arithmetic.Mean <- rowMeans(pm2.52000[,c('Arithmetic.Mean', 'Arithmetic.Mean2')], na.rm=TRUE)

pm2.52000 <- pm2.52000[c(1:4,6)]
pm2.52000 <- pm2.52000[order(pm2.52000$location, pm2.52000$Date.Local),]
names(pm2.52000)[names(pm2.52000) == "Date.Local"] <- "date"
names(pm2.52000)[names(pm2.52000) == "Arithmetic.Mean"] <- "pm2.5"
pm2.52000$pm2.5[pm2.52000$pm2.5<0] <-0
summary(pm2.52000)

pm2.5location <- unique(pm2.52000$location)
binder <- pm2.52000

binderF <- function(dataset1,dataset2, bind)
            {
              setwd("~/Desktop/historypm2.5")
              x <- gsub(" ", "", paste(as.character(dataset1), ".csv"))
              y <- gsub(" ", "", paste(as.character(dataset2), ".csv"))
              dataset1 <- read.csv(x)
              dataset2 <- read.csv(y)
              dataset1 <- dataset1[c(6,7,12,17,25,26)]
              dataset2 <- dataset2[c(6,7,12,17,25,26)]
              dataset1$Date.Local <- as.Date(dataset1$Date.Local)
              dataset2$Date.Local <- as.Date(dataset2$Date.Local)
              dataset1$location <-paste(as.character(dataset1$State.Name),as.character(dataset1$County.Name))
              dataset2$location <-paste(as.character(dataset2$State.Name),as.character(dataset2$County.Name))
              dataset1$Longitude <- as.integer(dataset1$Longitude)
              dataset1$Latitude <- as.integer(dataset1$Latitude)
              dataset2$Longitude <- as.integer(dataset2$Longitude)
              dataset2$Latitude <- as.integer(dataset2$Latitude)
              dataset1 <- dataset1[c(1,2,3,4,7)]
              dataset2 <- dataset2[c(1,2,3,4,7)]
              names(dataset2)[names(dataset2) == "Arithmetic.Mean"] <- "Arithmetic.Mean2"

              dataset1 <- merge(dataset1, dataset2,by = c("location","Date.Local","Longitude","Latitude"),all = TRUE)

              dataset1$Arithmetic.Mean <- rowMeans(dataset1[,c('Arithmetic.Mean', 'Arithmetic.Mean2')], na.rm=TRUE)

              dataset1 <- dataset1[c(1:5)]

              dataset1 <- dataset1[order(dataset1$location, dataset1$Date.Local),]
              names(dataset1)[names(dataset1) == "Date.Local"] <- "date"

              names(dataset1)[names(dataset1) == "Arithmetic.Mean"] <- "pm2.5"
              dataset1$pm2.5[dataset1$pm2.5<0] <-0
              bind <- rbind(bind,dataset1)
            
              bind <- bind[order(bind$location,bind$date),]
              return(bind)
            }

binder <- binderF("pm2.52001", "pm2.5n2001", binder )
binder <- binderF("pm2.52002", "pm2.5n2002", binder )
binder <- binderF("pm2.52003", "pm2.5n2003", binder )
binder <- binderF("pm2.52004", "pm2.5n2004", binder )
binder <- binderF("pm2.52005", "pm2.5n2005", binder )
binder <- binderF("pm2.52006", "pm2.5n2006", binder )
binder <- binderF("pm2.52007", "pm2.5n2007", binder )
binder <- binderF("pm2.52008", "pm2.5n2008", binder )
binder <- binderF("pm2.52009", "pm2.5n2009", binder )
binder <- binderF("pm2.52010", "pm2.5n2010", binder )
binder <- binderF("pm2.52011", "pm2.5n2011", binder )
binder <- binderF("pm2.52012", "pm2.5n2012", binder )
binder <- binderF("pm2.52013", "pm2.5n2013", binder )
binder <- binderF("pm2.52014", "pm2.5n2014", binder )
binder <- binderF("pm2.52015", "pm2.5n2015", binder )
binder <- binderF("pm2.52016", "pm2.5n2016", binder )
binder <- binderF("pm2.52017", "pm2.5n2017", binder )
binder <- binderF("pm2.52018", "pm2.5n2018", binder )
binder <- binderF("pm2.52019", "pm2.5n2019", binder )
binder <- binderF("pm2.52020", "pm2.5n2020", binder )
```

Now, we begin exploring the data.
```{r explorePM2.5, echo = FALSE}
nrow(binder)
summary(binder)
plot(table(binder$pm2.5))
head(binder)
```


We must now generate a table of the indicators of the PM2.5 longterm exposure.

We decided to compute the mean, median, standard deviation, maximum, minimum, and number of days from which these values were generated.We also created 3 indicators called normal, unhealthy, and hazardous. These three columns have the percentage of days that fall within the following ranges of PM2.5 air concentration:
  normal: below 12
  unhealthy: 13 to 35
  hazardous: above 35


```{r indicator, message=FALSE, warning=FALSE, out.width="33%", fig.show="hold", echo = FALSE}
#GENERATE INDICATOR TABLE
finalloc <- unique(binder$location)
indicator <- data.frame(finalloc)
indicator$mean <- 0
indicator$sd <- 0
indicator$median<- 0
indicator$var<- 0
indicator$min<- 0
indicator$max<- 0
indicator$normal<- 0
indicator$unhealthy<- 0
indicator$hazardous<- 0
indicator$accuracy <- 0
indicator$longitude <- 0
indicator$latitude <- 0
for (i in finalloc)
{
  x <- binder[binder$location == i,]
  indicator$mean[indicator$finalloc == i] <-  as.integer(mean(x$pm2.5, na.rm = FALSE))
  indicator$sd[indicator$finalloc == i] <- as.integer(sd(x$pm2.5))
  indicator$median[indicator$finalloc == i] <- as.integer(median(x$pm2.5))
  indicator$var[indicator$finalloc == i] <- as.integer(var(x$pm2.5))
  indicator$max[indicator$finalloc == i] <- as.integer(max(x$pm2.5))
  indicator$min[indicator$finalloc == i] <- as.integer(min(x$pm2.5))
  # 0-100, 101-200, 201- max
  indicator$normal[indicator$finalloc == i]<-  as.integer((sum(x$pm2.5 < 12)/length(x$pm2.5))*100)
  indicator$unhealthy[indicator$finalloc == i]<-  as.integer((sum(x$pm2.5 > 12 && x$pm2.5 < 35)/length(x$pm2.5))*100)
  indicator$hazardous[indicator$finalloc == i]<-  as.integer((sum(x$pm2.5 > 35)/length(x$pm2.5))*100)
  indicator$accuracy[indicator$finalloc == i] <-  as.integer((nrow(x))*100)
  indicator$longitude[indicator$finalloc == i] <- x$Longitude[1]
  indicator$latitude[indicator$finalloc == i] <- x$Latitude[1]
  
}


```

Now we look at the distribution of the computed indicators.
```{r exploreMore}
plot(table(indicator$mean))
plot(table(indicator$median))
plot(table(indicator$max))
plot(table(indicator$min))
plot(table(indicator$var))
plot(table(indicator$sd))
plot(table(indicator$normal))
plot(table(indicator$hazardous))
plot(table(indicator$unhealthy))
plot(table(indicator$accuracy))

# indicator$longitude
# setwd("C:/Users/MQAYM/Desktop") 
# write.csv(indicator,file="indicator.csv", row.names = FALSE)
```
Accordingly, we choose the 500 counties with the most reliable histories (most recorded number of days).

```{r}

location <- intersect(indicator$)

```


```{r}
setwd("~/Desktop")
cases2020 <- read.csv("casesdeaths.csv")
cases2020 <- cases2020[c(4,5,6,8,10,12,13)]
cases2020$location <-paste(as.character(cases2020$state),as.character(cases2020$location_name))
cases2020 <- distinct(cases2020[c(3:8)])
cases2020$date <- as.Date(cases2020$date)
names(cases2020)[names(cases2020) == "date"] <- "Date.Local"
cases2020 <- cases2020[order(cases2020$location,cases2020$Date.Local),]
summary(cases2020)
# new_cases and new_deaths have negative values
cases2020$new_cases[cases2020$new_cases < 0] <- NA
cases2020$new_deaths[cases2020$new_deaths < 0] <- NA
summary(cases2020)
newCases <- cases2020[1,]

finalloc <- unique(intersect(cases2020$location, binder$location))
for (i in finalloc)
{
  x <- cases2020[cases2020$location == i,]
  x[1,]$new_cases <- 0
  x[1,]$new_deaths <- 0
  for (j in 2:nrow(x))
  {
    if (x[j,]$cumulative_cases < x[j-1,]$cumulative_cases)
      {
      
      x[j,]$cumulative_cases <- x[j-1,]$cumulative_cases
      }
    if (x[j,]$cumulative_deaths < x[j-1,]$cumulative_deaths)
    {
      x[j,]$cumulative_deaths <- x[j-1,]$cumulative_deaths
      }
    if (is.na(x[j,]$new_cases)) { x[j,]$new_cases <- x[j,]$cumulative_cases - x[j-1,]$cumulative_cases }
    if (is.na(x[j,]$new_deaths)) { x[j,]$new_deaths <- x[j,]$cumulative_deaths - x[j-1,]$cumulative_deaths }
  }
  newCases <- rbind(newCases,x)
}

newCases <- newCases[2:nrow(newCases),]
summary(newCases)
write.csv(newCases,file = "newCases.csv")
```

```{r bindCasesIndicator}

setwd("~/Desktop")
health <- read.csv("final_vars.csv")
# indicator <- read.csv("indicator.csv")
health <- health[c(1:27,33)]
head(health)
names(indicator)[names(indicator) == "finalloc"] <- "location"
vars <- merge(indicator,health, by ="location")
summary(vars)

```


```{r tempplot, echo=FALSE, fig.show="hold", message=FALSE, warning=FALSE, out.width="25%"}
setwd("~/Desktop/temp")
temp2020 <- read.csv("temp2020.csv")
temp2020 <- temp2020[c(12,17,25,26)]
temp2020$location <-paste(as.character(temp2020$State.Name),as.character(temp2020$County.Name))
temp2020 <- distinct(temp2020[c(1,2,5)])
temp2020$Date.Local <- as.Date(temp2020$Date.Local)
names(temp2020)[names(temp2020) == "Arithmetic.Mean"] <- "temp"
temp2020<-aggregate(x = temp2020$temp, by = list(temp2020$Date.Local,temp2020$location), FUN = mean)
names(temp2020)[names(temp2020) == "x"] <- "temp"
names(temp2020)[names(temp2020) == "Group.1"] <- "Date.Local"
names(temp2020)[names(temp2020) == "Group.2"] <- "location"
uniqueL <- unique(temp2020$location)
sz <- length(uniqueL)
for (i in 1:20)
{
  x <- sample(2:sz, 1)
  sm <- temp2020[temp2020$location==uniqueL[x],]
  print(ggplot(sm, aes(Date.Local,temp))+geom_line()+geom_smooth(span = 0.3) +ggtitle(uniqueL[x]))
}
```

``` {r wind, message=FALSE, warning=FALSE, out.width="25%", fig.show="hold", echo = FALSE}
#GETTING WIND2020
setwd("~/Desktop/wind")
wind2020 <- read.csv("wind2020.csv")
wind2020 <- wind2020[c(12,17,25,26)]
wind2020$location <-paste(as.character(wind2020$State.Name),as.character(wind2020$County.Name))
wind2020 <- distinct(wind2020[c(1,2,5)])
wind2020 <- wind2020[order(wind2020$location,wind2020$Date.Local),]
names(wind2020)[names(wind2020) == "Arithmetic.Mean"] <- "wind"
wind2020<-aggregate(x = wind2020$wind, by = list(wind2020$Date.Local,wind2020$location), FUN = mean)
names(wind2020)[names(wind2020) == "x"] <- "wind"
names(wind2020)[names(wind2020) == "Group.1"] <- "Date.Local"
names(wind2020)[names(wind2020) == "Group.2"] <- "location"
wind2020$Date.Local <- as.Date(wind2020$Date.Local)
uniqueL <- unique(wind2020$location)
sz <- length(uniqueL)
for (i in 1:20)
{
  x <- sample(2:sz, 1)
  sm <- wind2020[wind2020$location==uniqueL[x],]
  print(ggplot(sm, aes(Date.Local,wind))+geom_line() +geom_smooth(span = 0.3) +ggtitle(uniqueL[x]))
}
```
```{r}

```
```{r locations}
locations <- intersect(vars$location, intersect(cases2020$location,intersect(wind2020$location,temp2020$location)))
vars <- vars[vars$location %in% locations,]
vars <- vars[vars$accuracy > quantile(vars$accuracy, c(0.25)),]
FINAL_LOCATIONS <- vars$location
temp2020 <- temp2020[temp2020$location %in% FINAL_LOCATIONS,]
wind2020 <-wind2020[wind2020$location %in% FINAL_LOCATIONS,]
cases2020$Date.Local <- as.Date(cases2020$Date.Local )
cases2020 <- cases2020[cases2020$location %in% FINAL_LOCATIONS,]
daily_data <- merge(cases2020,merge(wind2020,temp2020))
daily_data <- daily_data[order(daily_data$location,daily_data$location),]

non_daily_data <- vars
data <- merge(non_daily_data,cases2020)

#Creating Lags
library(DataCombine)
locations <- unique(data$location)

lag1 <- data.frame()
for (i in locations){
  print("here1")
  l <- subset(data, data$location==i)
  print(l)
  ndeaths<- l$new_deaths
  print(ndeaths)
  print("here3")
  for (j in c(1:10)){
    print(ndeaths)
    ndeaths <- cbind(ndeaths,shift(ndeaths,j))
  }
  print(ndeaths)
  lag1 <- cbind(lag1,ndeaths)
}

lag2 <- data.frame()
for (i in locations){
  l <- subset(data, data$location==i)
  ncases <- l$new_cases
  for (j in c(1:9)){
    ncases <- cbind(ncases,shift(ncases,j))
  }
  lag2 <- rbind(lag2,ncases)
}
lag3 <- data.frame()
for (i in locations){
  l <- subset(data, data$location==i)
  wind <- l$wind
  for (j in c(1:9)){
    wind <- cbind(wind,shift(wind,j))
  }
  lag3 <- rbind(lag3,wind)
}
lag4 <- data.frame()
for (i in locations){
  l <- subset(data, data$location==i)
  temp <- l$temp
  for (j in c(1:9)){
    temp <- cbind(temp,shift(temp,j))
  }
  lag4 <- rbind(lag4,wind)
}


summary(data)
data <- cbind(data,cbind(lag2,lag))
```

```{r}
write.csv(vars$location, "locFinal.csv", row.names = FALSE)

plot(table(vars$mean))
plot(table(vars$median))
plot(table(vars$sd))
plot(table(vars$hazardous))

```



